---
title: "sankey diagram for visualizing clinical pathways for severe sepsis patients"
author: "Xing Song"
date: "August 8, 2018"
output: 
  html_document:
    keep_md: false 
---

## Instructions for reading the Sankey diagram:

* **Acronyms**:
    + SI: suspect infection onset
    + SIRS2: two SIRS onset
    + OD1: first site of organ dysfunctio onset
    + AC: abnormal coagulation
    + KF: kidney failure
    + IL: increased lactate
    + AMS: altered mental status
    + LF: liver failure
    + BLOOD_C: first blood cultuer order
    + ABX: first antibiotic administration
    
***

* **Temporal indicator**:
    + "@X" inidicates the relative temporal order of an event along an existing pathway
    
***
  
* **Color Codes**:
    + pink: sepsis alerts
    + blue: treatment bundle components
    + green: a pathway with lower risk (empirical odd ration <= 1) to sepsis onset; the darker the green is the lower the risk is 
    + red: a pathway with higher risk (empirical odd ratio > 1) to sepsis onset; the darker the green is the lower the risk is
    + grey: a pathway lead to no sepsis within 48 hours since triage
    
***

* **Tooltip Messages**:
    + if you hover over each node, you can see exactly how many patient-encounters have reached that state
    + if you hover over each link, you can see exactly how many patient-encounters have experienced that transition and the common time lapse for that transition
      ([t1, t2, t3] are corresponding to time at [25th percentile, 50th percentile, 75th percentile])
    + size of links and nodes is proportional to patient-encounter size
    
***
```{r set,include=F}
source("./R/util.R")
require_libraries(c("networkD3",
                    "tidyr",
                    "dplyr",
                    "magrittr",
                    "leaflet"))
```


```{r echo=FALSE}
#load data
sankey_dat<-readRDS("./data/sankey_dat.rda")

#customize link and node colors
my_color<-'d3.scaleOrdinal() 
          .domain(["sympt","both","0","1","2","3","4","5"]) 
          .range(["#ffd7ff", "#5fd7ff","#4eaa5f","#99db7f","#f1ff75","#f9956d","#f77676","#e04e4e"])'

#generate basic plot (not print)
sn<-networkD3::sankeyNetwork(Links = sankey_dat$links,
                             Nodes = sankey_dat$nodes,
                             Source = "source",
                             Target = "target",
                             Value = "value",
                             NodeID = "key",
                             colourScale = my_color,
                             LinkGroup = "group",
                             NodeGroup = "group",
                             fontSize= 15, nodeWidth = 30,
                             width = 2000,height=1500)

#add more info to links
links_name<-sankey_dat$links %>%
  inner_join(sankey_dat$nodes %>% dplyr::select(key) %>% mutate(code=1:n()-1),
             by=c("source"="code")) %>%
  dplyr::rename(source_name = key) %>%
  inner_join(sankey_dat$nodes %>% dplyr::select(key) %>% mutate(code=1:n()-1),
             by=c("target"="code")) %>%
  dplyr::rename(target_name = key)

#customize tooltip label
sn$x$links$label<-paste(paste(links_name$source_name,links_name$target_name,sep="->"),
                        paste0(sankey_dat$links$value,sankey_dat$links$label),sep="\n")
htmlwidgets::onRender(
  sn,
  '
  //user-defined label
  function(el,x){
  d3.selectAll(".link")
  .select("title")
  .text(function(d) { return d.label; });
  
  //move position of node text
  d3.select(el)
  .selectAll(".node text")
  .attr("text-anchor","begin")
  .attr("x",5)
  }
  '
)
```



